### 2022-10-31 files/Archive/NRF.tex
initial commit

\documentclass[../../main.tex]{subfiles}
\begin{document}
\setcounter{chapter}{1}
\setcounter{section}{0}
\section{Introduction}

\textcolor{red}{**********Sit in introduction*************}

It is possible to derive a system of partial differential equations with corresponding boundary conditions but due to the complexity of the system, a different approach is considered. Consider the model problem in variational form.

We consider the {\bf variational form} of the vibration problem. This variational problem may be used to implement the {\bf finite element method}. The finite element method can be used to simulate vibrations of the elastic body for both dynamic and static problems and also calculate the natural frequencies and modes of the elastic body. Existence and uniqueness of solutions to the problem can also be investigated using a weak version of the variational form.

\subsubsection*{A Dynamic Problem}
Find a vector valued function $u \in C(\Omega)$ such that it satisfies the equations of motion and the constitutive equations. Since $u \in C(\Omega)$, $u$ satisfies the boundary conditions defined on $\Sigma$ and $\Gamma$.\\

Initial conditions for $u$ will need to be defined.

\subsubsection*{A Static Problem}
The static porblem is a special case of the dynamic problem  where $\partial_t^2 u = 0$. In this case, there are also no initial conditions.\\
\textcolor{red}{**************************}

\textcolor{red}{Research into the behaviour of suspension bridges is still a widely researched field by both engineers and mathematicians. This research started as several bridges suffered collapse. }

\subsubsection{Different Research Approaches}
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}
		\draw (3,2) -- (7,2);
		\draw (3,2) -- (3,1.9);
		\draw (7,2) -- (7,1.9);
		\node at (5,2.1) {Main Span};

		\draw[line width = 0.25mm] (0,0) -- (10,0);

		\draw[line width = 0.5mm] (1.4,0) -- (1.4,-0.5);
		\draw[line width = 2mm] (1.4,-0.5) -- (1.4,-0.7);

		\draw[line width = 0.5mm] (2.2,0) -- (2.2,-0.5);
		\draw[line width = 2mm] (2.2,-0.5) -- (2.2,-0.7);

		\draw[line width = 0.5mm] (8.6,0) -- (8.6,-0.5);
		\draw[line width = 2mm] (8.6,-0.5) -- (8.6,-0.7);

		\draw[line width = 0.5mm] (7.8,0) -- (7.8,-0.5);
		\draw[line width = 2mm] (7.8,-0.5) -- (7.8,-0.7);

		\draw[line width = 0.5mm] (3,1.5) -- (3,-0.5);
		\draw[line width = 2mm] (3,-0.5) -- (3,-0.7);

		\draw[line width = 0.5mm] (7,1.5) -- (7,-0.5);
		\draw[line width = 2mm] (7,-0.5) -- (7,-0.7);

		\draw (5,0.4) parabola (3,1.5);
		\draw (5,0.4) parabola (7,1.5) ;

		\draw (3,1.5) -- (0.2,-0.5);
		\draw (7,1.5) -- (9.8,-0.5);

		\draw[line width = 2mm] (0.2,-0.5) -- (0.2,-0.7);
		\draw[line width = 2mm] (9.8,-0.5) -- (9.8,-0.7);

		\draw[line width = 0.1mm] (5,0) -- (5,0.4);
		\draw[line width = 0.1mm] (5.25,0) -- (5.25,0.43);
		\draw[line width = 0.1mm] (5.5,0) -- (5.5,0.46);
		\draw[line width = 0.1mm] (5.75,0) -- (5.75,0.55);
		\draw[line width = 0.1mm] (6,0) -- (6,0.69);
		\draw[line width = 0.1mm] (6.25,0) -- (6.25,0.85);
		\draw[line width = 0.1mm] (6.5,0) -- (6.5,1);
		\draw[line width = 0.1mm] (6.75,0) -- (6.75,1.25);

		\draw[line width = 0.1mm] (4.75,0) -- (4.75,0.43);
		\draw[line width = 0.1mm] (4.5,0) -- (4.5,0.46);
		\draw[line width = 0.1mm] (4.25,0) -- (4.25,0.55);
		\draw[line width = 0.1mm] (4,0) -- (4,0.69);
		\draw[line width = 0.1mm] (3.75,0) -- (3.75,0.85);
		\draw[line width = 0.1mm] (3.5,0) -- (3.5,1);
		\draw[line width = 0.1mm] (3.25,0) -- (3.25,1.25);
	\end{tikzpicture}
	\caption{Side view of single span suspension bridge with conventional bridge on either sides of the main span.}
\end{figure}

\subsubsection{Motivation}

In most real-world applications, three-dimensional models are a natural choice. But it may not be the most efficient or practical choice. When considering the three-dimensional models, there is a limitation on the availability of computing power that is required for these models as well as the complexity of the models. This is especially true of systems of multiple three-dimensional models. Simplifications of the models are a great way around this limitation.\\

In many applications, especially engineering, one-dimensional models such as the Timoshenko beam model is used to a great extent for most beam type applications. To do this, some assumptions about the properties of the beam are required to be made. This brings into question the validity of using a one-dimensional model over a more realistic three-dimensional model.\\

In this essay, we consider three different models for a beam. A three-dimensional elastic solid is compared to a one-dimensional beam model, namely the Timoshenko beam model. As a natural intermediate step, we will also look at a two-dimensional elastic solid derived using an assumption of plane stress.\\

The comparisons of the models are made for a cantilever beam. The comparisons between the models are made by considering the modes and natural frequencies of the models. This is a valid comparison since the solution of a vibration problems can be written in terms of a linear combination of the eigenfunctions.\\

In Chapter 1, the models for a three-dimensional elastic solid, a two dimensional elastic solid as well as a one-dimensional Timoshenko beam model is presented. In Chapter 2, we look at the exsistance and uniqueness of solutions to these linear vibration models. Chapter 3 covers the convergence of the eigenvalue problems as well as a dynamic problem. This is neccessary since we intend to use the Finite Element Method in Chapter 5 to derive and solve a static problem and the eigenvalue problem for the two and three-dimensional models. For the one-dimensional Timoshenko model, seperation of variables can be applied to obtain the eigenvalues exactly and this is done in Chapter 4. In Chapter 4, we also look at the results of an article by [] that tested the validity of the Timoshenko beam model using an empirical test.

\textcolor{red}{************************************************}\\
\subsubsection*{Research rationale and motivation}
From small transverse vibration of a linearly elastic (flat) plate [Mindlin, 1951], to a fully non-linear shell model, numerous mathematical problems are possible. (The same remark is also true for beams.) In this research project, our focus is on the small vibrations but not sufficiently small to justify a linear model.\\

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\setcounter{chapter}{1}
\setcounter{section}{1}
\section{NRF Research Proposal}
\subsection*{Finite Element Analysis of Model Problems for Elastic Rods and Shells}
\subsubsection*{Research rationale and motivation}
From small transverse vibration of a linearly elastic (flat) plate [Mindlin, 1951], to a fully non-linear shell model, numerous mathematical problems are possible. (The same remark is also true for beams.) In this research project, our focus is on the small vibrations but not sufficiently small to justify a linear model.\\

### 2022-10-31 files/Chapter7/Beam.tex
initial commit

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of the Timoshenko beam model}
This section is a literature study of the article \cite{LVV09}. The goal of the authors is to show that for practical applications, the solution of the Timoshenko beam model is close to the solution of a two-dimensional beam model. This is referred to as the validity of the Timoshenko beam model in this dissertation.\\

For practical applications, it would be preferred to compare the Timoshenko beam model to a more realistic three-dimensional model. But the authors of \cite{LVV09} explain that this will introduce complexities. \textcolor{red}{What complexities?} To mitigate some of the complexities, the two-dimensional model is used as an intermediate step.\\

Where \cite{LVV09} differs from most other research, is that the boundary conditions are also taken into consideration. The authors of \cite{LVV09} use a cantilever beam configuration for their investigation.\\

To compare the two models, the authors compare the natural frequencies of the models. From modal analysis, it is known that a solution of a second order hyperbolic type problem is a linear combination of the eigenvalues and eigenfunctions of the model. Therefore it is sufficient to compare the eigenvalues and eigenfunctions to determine the validity of the different dimensional models.\\

\subsection{The models}
Consider the cantilever Timoshenko beam model problem Problem T-2 from section \ref{ssec:1D_Model:ModelProblems} and the cantilever two-dimensional beam model problem Problem 2D-1 from section \ref{ssec:2D_Model:ModelProblem}.\\

Figure 6.1 shows the two beams side-by-side for a visual guide.

@@

The parameter $h$ is the height of the two-dimensional beam. Recall the dimensionless constant $\displaystyle \alpha = \frac{A \ell^2}{I}$ from section \ref{sssec:1D_Model:DimensionlessForm}. For the Timoshenko beam model with a rectangular cross-section, the area moment of inertial is 
\begin{eqnarray*}
	I = \frac{h^3b}{12}.
\end{eqnarray*}

After substitution, the following relation ship between $\alpha$ and $h$ is found
\begin{eqnarray*}
	h = \sqrt{\frac{12}{\alpha}}.
\end{eqnarray*}
The parameter $\alpha$ can be chosen in such a way that the beams have equal dimensions.

\subsection{Accuracy of the numerical eigenvalues}
For the eigenvalues and eigenfunctions of the Timoshenko beam, the authors of \cite{LVV09} reference an another article \cite{VV06}. In this article \cite{VV06}, a method is described to obtain the eigenvalues and eigenfunctions of a (cantilever) Timoshenko beam. This method also allows for the eigenvalues to be obtained to a desired level of accuracy (using a numerical method such as interval division). In this dissertation, \cite{VV06} is discussed in section \ref{sec:Timo:EigenvalueProblem}.\\

For the two-dimensional beam, the Finite Element Method is used to calculate the eigenvalues and eigenfunctions. In \cite{LVV09} the authors used piecewise bi-cubic basis functions and a grid size large enough to ensure that the eigenvalues are accurate to 3 significant digits.\\

@@

\subsection{Comparing the shape of the eigenfunctions}
Before the eigenvalues of the beams can be compared, a method is required to be able to match the eigenvalues. In \cite{LVV09} the authors compare the modal shapes of the Timoshenko beam model to the modal shapes of the two-dimensional model.\\

Similar modal shapes implies that the eigenvalues can be matched up and compared. Eigenvalues relating to the shapes of the Timoshenko beams modes are called beam-type eigenvalues by \cite{LLV09}.

\subsubsection{Shapes relating to beam-type eigenvalues}
The following are examples of beam-type mode shapes for the displacement $w$ for the cantilever two-dimensional beam and cantilever Timoshenko beam. The figures are presented side-by-side.

@@

\subsubsection{Shapes relating to non-beam type eigenvalues}
The following are examples of non beam-type mode shapes for the displacement $w$ of the cantilever two-dimensional beam.
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam1.jpg}
				\subcaption{Non-Beam Type - $\lambda_4 = 7.7077$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam2.jpg}
				\subcaption{Non-Beam Type - $\lambda_8 = 69.344$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the non-beam type 2D body with $h = 1/20$.}
		}
	}
\end{figure}
\FloatBarrier

Using this method all the relevant eigenvalues in this section were matched. The non-beam type eigenvalues are highlighted in gray in the tables.

\subsubsection{Shapes of the cross-section}
The Timoshenko beam theory differs from other one-dimensional beam theories by also including shear. It is assumed that the cross-sections need not remain perpendicular to the neutral axis of the beam. The cross-sections remain straight.\\

For the two-dimensional body, the shape of the cross-section of beam type eigenfunctions can deform into an a s-shape. Figure \ref{fig:s} shows an exaggerated representation of the s-shaped deformation. This is also given and explained in \cite{LVV09}.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=1.2]

@@

\end{tikzpicture}
	\caption{S-shape deformation and average rotation.}\label{fig:s}
\end{figure}
The red line in figure \ref{fig:s} show how the average rotation of the cross-section is calculated in \cite{LVV09}.
\FloatBarrier

\subsubsection{Direct comparison of mode shapes}
Figure \ref{fig:w} shows the displacement $w$ of the cantilever Timoshenko beam and the cantilever two-dimensional beam directly compared.

@@

Similarly Figure \ref{fig:phi} shows the direct comparison of the angle of the cross-section of the cantilever Timoshenko beam and the two-dimensional beam. The average rotation of the cross-section of the two-dimensional beam's mode shape is calculated as shown in figure \ref{fig:s}.

@@

Figures \ref{fig:w} and \ref{fig:phi} show that for $\alpha = 4800$, the mode shapes are very close, even for the eigenvalues 8 (for Timoshenko beam) and 10 (for the two-dimensional beam). The authors of \cite{LVV09} obtained similar results for $\alpha = 300$ for the eigenvalues 6 (for Timoshenko beam) and 9 (for the two-dimensional beam). Therefore a direct comparison of the mode shapes are feasible for short and long beams.

\subsubsection{Remark}
Due to the fact that any multiple of an eigenfunction is still an eigenfunction, the figures can be scaled. This is why no scale is indicated on the y-axis in \cite{LVV09}. Figures \ref{fig:w} and \ref{fig:phi} were scaled by eye.

@@

\section{Validity of the two-dimensional elastic model}
This section continues the work of the article \cite{LVV09}. As mentioned by the authors of \cite{LVV09}, a comparison using a three-dimensional beam is desired. The two-dimensional beam model is an intermediate step to be able to investigate the validity of the Timoshenko beam model using a three-dimensional beam model as a reference. As mentioned by the authors, this will avoid some complexities.\\

The approach for this section remain similar to the previous section and the work of \cite{LVV09}. The eigenvalues for both models are calculated using FEM. The eigenvalues are then matched by sketching and comparing the mode shapes. Similar shapes indicate the the eigenvalues are comparable. The eigenvalues are then compared to determine how close the solutions of the models are.\\

As seen in the previous section, the two-dimensional model has eigenvalues not relating to beam type eigenvalues. There will also be eigenvalues not relating to beam type eigenvalues in this section. The results will include some of these eigenvalues, but the focus of our investigation remain with beam problems.

\subsection{The models}
Consider a cantilever two-dimensional beam, referred to as Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem} and a cantilever three-dimensional beam, referred to as Problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.\\

@@

@@

\subsection{Comparing the mode shapes}
As in \cite{LVV09}, the shape of the modes are compared. Similar mode shapes implies that the eigenvalues are comparable. Results for both beam and non-beam type eigenvalues are included. The focus still remains on beam type eigenvalues.

\subsubsection{Mode shapes relating to beam type eigenvalues.}
Figure \ref{fig:beam-2dv3d} show some examples of beam type mode shapes for the displacement $u$. These are shapes that can be obtained from the Timoshenko beam model.

@@

\subsubsection{Mode shapes relating to non-beam type eigenvalues that are present in the two-dimensional model.}
Figure \ref{fig:nonbeam-2dv3d} show examples of mode shapes relating to non-beam type eigenvalues for the displacement $u$. These mode shapes are the non-beam type eigenvalues found in both the two-dimensional beam and the three-dimensional beam model.

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Validity of the Timoshenko beam model}
This section is a literature study of the article \cite{LVV09}. The goal of \cite{LVV09} to show that the one-dimensional Timoshenko beam theory, is a feasible model to use in practical applications of beams. In the real world, a beam is a three-dimensional model. Therefore it is important to validate the use of the Timoshenko beam theory in practical applications.\\

In mathematical terms, the validity of the Timoshenko beam theory is described as how well the solutions of the Timoshenko beam theory compare to more realistic, higher-dimensional models. It is obvious that a three-dimensional beam model is a better approximation of a real world beam. But the Timoshenko beam theory is less complex and more widely used.\\

The authors of \cite{LVV09} use a two-dimensional cantilever beam, to validate the use of a cantilever Timoshenko beam for practical importance. Although a direct comparison to a three-dimensional beam would be preferred, the authors state that this would introduce complexities. A two-dimensional model is therefore used as an intermediate step.\\

Where \cite{LVV09} differs from most other research, is that the boundary conditions are also taken into consideration.\\

To compare the two models, the authors compare the natural frequencies of the models. From modal analysis, it is known that a solution of a second order hyperbolic type problem is a linear combination of the eigenvalues and eigenfunctions of the model. Therefore it is sufficient to compare the eigenvalues and eigenfunctions to compare the solutions and comment on the validity of the models.\\

\subsection{The models}
Recall the cantilever Timoshenko beam model from section \ref{ssec:1D_Model:ModelProblems}, referred to as Problem T-2 and the cantilever two-dimensional beam model from section \ref{ssec:2D_Model:ModelProblem}, referred to as Problem 2D-1. In \cite{LVV09} it is assumed that the beams have a rectangular cross-section.\\

Figure 6.1 shows the two beams side-by-side.

@@

In Chapter 1, the following parameters were established. The parameter $h$ is the height of the two-dimensional beam and for the Timoshenko dimensionless constant $\displaystyle \alpha = \frac{A \ell^2}{I}$ was found.\\

For the Timoshenko beam model with a rectangular cross-section, the area moment of inertial is calculated as
\begin{eqnarray*}
	I = \frac{h^3b}{12}.
\end{eqnarray*}

After substitution, the following relationship between $\alpha$ and $h$ is found
\begin{eqnarray*}
	h = \sqrt{\frac{12}{\alpha}}.
\end{eqnarray*}
Using this relationship, $\alpha$ can be chosen in such a way that the beams have equal lateral dimensions.

\subsection{Accuracy of the numerical eigenvalues}
The calculate the eigenvalues and eigenfunctions of the Timoshenko beam, the authors of \cite{LVV09} use the article \cite{VV06}. This article was discussed in section \ref{sec:Timo:EigenvalueProblem} in this dissertation. It is important to note that the method described in \cite{VV06} allows us to find the eigenvalues of a cantilever Timoshenko beam accurate to a desired level, using numerical methods like interval division.\\

A different approach is needed for the two-dimensional beam. In section \ref{sec:FEM:2D}, the Finite Element Method was used to obtain a numerical formula to solve the eigenvalue problem Problem 2D-1E. In \cite{LVV09} the authors used piecewise bi-cubic basis functions and a grid size large enough to ensure that the eigenvalues are accurate to 3 significant digits.\\

@@

\subsection{Comparing the shape of the eigenfunctions}
Before the eigenvalues can be compared, a method is required to be able to match the eigenvalues of the two different models. The authors of \cite{LVV09} match the modal shapes of the eigenfunctions. The corresponding eigenvalues can then be matched.\\

The eigenvalues of the Timoshenko beam, are referred to by the authors as beam-type eigenvalues.

\subsubsection{Shapes relating to beam-type eigenvalues}
The following figures are examples of beam-type mode shapes for the displacement $w$.

@@

\subsubsection{Shapes relating to non-beam type eigenvalues}
The following figures are examples of non beam-type mode shapes for the displacement $w$. These mode shapes are not present in the cantilever Timoshenko beam model and are not beam related. 
\begin{figure}[h!]
	\scalebox{.8}{
		\makebox[\textwidth][c]{
			\centering
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam1.jpg}
				\subcaption{Non-Beam Type - $\lambda_4 = 7.7077$}
				\label{fig:minipage2}
			\end{minipage}
			\begin{minipage}[b]{0.8\linewidth}
				\includegraphics[width=1\linewidth]{NotBeam2.jpg}
				\subcaption{Non-Beam Type - $\lambda_8 = 69.344$}
				\label{fig:minipage1}
			\end{minipage}
			\caption{Modal shapes of the displacement $w$ for the non-beam type 2D body with $h = 1/20$.}
		}
	}
\end{figure}
\FloatBarrier

The existence of these non-beam type eigenvalues imply that there are eigenvalues in the two-dimensional model, that are not in the Timoshenko beam model. This is an example of the complexities that arise when comparing models of different dimensions.

\subsubsection{Shapes of the cross-section}
The Timoshenko beam theory improve some one-dimensional beam theories such as the Euler-Bernoulli beam theory by also including the effect of shear. In the Timoshenko beam theory it is assumed that the cross-sections need not remain perpendicular to the neutral axis of the beam. The cross-sections however remain a straight line.\\

For the two-dimensional beam, the shape of the cross-section can deform into a s-shape. This is explained in \cite{LVV09} and a similar figure in \cite{LVV09} is given here.
\FloatBarrier
\begin{figure}[h!]
	\centering
	\begin{tikzpicture}[scale=1.2]

@@

\end{tikzpicture}
	\caption{S-shape deformation of the cross-section.}\label{fig:s}
\end{figure}

Note that figure \ref{fig:s} shows an exaggerated example of the deformation of a cross-section. The red line shows how the average rotation of a cross-section is calculated in \cite{LVV09}. This average rotation can be used to compare the rotation of the cross-section of the two-dimensional beam to the rotation of the cross-section of the Timoshenko beam.
\FloatBarrier

\subsubsection{Direct comparison of mode shapes}
Figure \ref{fig:w} shows the displacement of the material lines of the cantilever Timoshenko beam and the cantilever two-dimensional beam, compared directly.

@@

Similarly Figure \ref{fig:phi} shows the direct comparison of the angle of the cross-section of the cantilever Timoshenko beam and the two-dimensional beam. The average rotation of the cross-section of the two-dimensional beam's mode shape is calculated as shown in figure \ref{fig:s} and \cite{LVV09}.

@@

In \cite{LVV09} the authors obtained similar results for $\alpha=300$, which represents a short beam. Both figures \ref{fig:w} and \ref{fig:phi} were calculated with5 $\alpha = 4800$, which represents a longer, narrower beam.

\subsubsection{Remark}
The mode shapes were scaled to obtain figures \ref{fig:w} and \ref{fig:phi}. Therefore no scale is shown on the y-axis.

@@

\section{Validity of the two-dimensional elastic model}
In this section, the validity of a two-dimensional cantilever beam is investigated. A three-dimensional cantilever beam is the model we'll use the measure. Recall that the two-dimensional model is an intermediate step between a Timoshenko beam and a three-dimensional beam.\\

In Chapter 1, models for a two-dimensional cantilever elastic body, as well as a three-dimensional cantilever elastic body is given. In Chapter 5, the Finite Element Method is applied to these models to obtain a formula to numerically calculate the eigenvalues and eigenfunctions of each of the models.\\

This section follows the same strategy as the previous section by comparing the eigenvalues and eigenfunctions of the models to determine the validity of the two-dimensional model. This investigation remains focussed on beam type problems.

\subsection{The models}
Recall a cantilever two-dimensional beam, referred to as Problem 2D-1 in section \ref{ssec:2D_Model:ModelProblem} and a cantilever three-dimensional beam, referred to as Problem 3D-1 in section \ref{ssec:3D_Model:ModelProblems}.\\

@@

For different values of $h$ and $b$, the first 30 eigenvalues are calculated with increasing grid sizes. The maximum relative error between the sets of eigenvalues are then calculated and graphed.\\

@@

\subsection{Comparing the mode shapes}
In this section, the mode shapes of the two dimensional and three-dimensional beam are compared. Recall that beam-type eigenvalues are defined by \cite{LVV09} as any eigenvalue for which it's corresponding mode shape is similar to the mode shapes obtained in the Timoshenko beam theory.

\subsubsection{Mode shapes relating to beam type eigenvalues.}
Figure \ref{fig:beam-2dv3d} show some examples of beam type mode shapes for the displacement $u$.

@@

\subsubsection{Mode shapes relating to non-beam type eigenvalues that are present in the two-dimensional model.}
Figure \ref{fig:nonbeam-2dv3d} show examples of mode shapes relating to non-beam type eigenvalues for the displacement $u$.
